---
title: "Lab .mono[001]"
subtitle: "Data cleaning and workflow [2/N]"
author: "Edward Rubin"
#date: "`r format(Sys.time(), '%d %B %Y')`"
date: "15 January 2021"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    # self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---
exclude: true

```{R, setup, include = F}
library(pacman)
p_load(
  broom, tidyverse,
  ggplot2, ggthemes, ggforce, ggridges, cowplot,
  latex2exp, viridis, extrafont, gridExtra, plotly,
  kableExtra, snakecase, janitor, DT,
  data.table, dplyr,
  lubridate, knitr, future, furrr,
  estimatr, FNN, caret, parsnip,
  huxtable, here, magrittr, tidyimpute
)
# Define colors
red_pink   = "#e64173"
turquoise  = "#20B2AA"
orange     = "#FFA500"
red        = "#fb6107"
blue       = "#3b3b9a"
green      = "#8bb174"
grey_light = "grey70"
grey_mid   = "grey50"
grey_dark  = "grey20"
purple     = "#6A5ACD"
slate      = "#314f4f"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  warning = F,
  message = F
)
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(knitr.table.format = "html")
```

---
layout: true
# .mono[mutate]

---
class: inverse, middle
---
name: mutate-refresh
## Refresher

You already know some of .b[`mutate()`]

- .b.pink[Create] new variables
- .b.blue[Transform] existing data into more usable forms
- .b.orange[Fill in] missing values

...
--
 but it turns out .b[there's more!]

---
name: pipe-refresh
## Pipes

But first, a reminder about pipes and the `magrittr` package.

The `%>%` operator is called a .hi[pipe].
```{r, eval = F}
# Equivalent:
some_function(a, b)
a `%>%` some_function(b)
```

--

The `%<>%` operator is an .b.purple[assignment pipe]. 
```{r, eval = F}
# Equivalent:
house_df = house_df %>% mutate(n_rooms = bed + bath + other)
house_df `%<>%` mutate(n_rooms = bed + bath + other)
```

---
name: sl-classes
## 'Regular' .mono[mutate]

--

Transform a numeric variable...

```{r, eval = F}
# Convert kg to lbs: 1 kg is ~2.205 pounds
starwars %<>% mutate(weight = mass * 2.205) 
# Check results
starwars %>% select(name, mass, weight)
```

--

```{r, echo = F}
# Convert kg to lbs: 1 kg is ~2.205 pounds
starwars %<>% mutate(weight = mass * 2.205) 
# Check results
starwars %>% select(name, mass, weight) %>% datatable(rownames = F, options = list(dom = 't', pageLength = 3))
```

`mutate()` receives data and returns both old and new columns. 
--
<br>`transmute()` returns only the new columns.

---
layout: true
# .mono[mutate] and family
---
class: inverse, middle
---

name: sl-classes
## Multiple columns

To `mutate()` multiple variables, you'll want one of its cousins:

- `mutate_at()` applies your function(s) to all .b.pink[specified] columns.

- `mutate_if()` applies your function(s) to .b.orange["matched"] columns.

- `mutate_all()` applies your function(s) to .b.slate[all] columns.

The main difference between the `mutate_*()` functions: 
> How do you want to select your columns?

.attn[Note:] These `mutate_*()` functions will .b.purple[replace] variables.

---
layout: true

# .mono[mutate] and family
## .mono[\_at]

`mutate_at()` wants three arguments:

1. .b.pink[`.tbl`] the dataset you're trying to mutate/transform (*e.g.*, `starwars`)
1. .b.orange[`.vars`] the specific variables you want to transform (*e.g.*, `mass`)
1. .b.purple[`.funs`] the function(s) you are using for the transformation (*e.g.*, `log`)

---
name: mutate-at-intro

--

For example, to take the `log` of `height` and `mass`

```{r, eval = F}
# Option 1: Character vector of variable names
starwars %>% mutate_at(.vars = `c("height", "mass")`, .funs = log)
```

.note[Above:] You can specify .orange[which variables] using a character vector.

--

.attn[Q] Where is .pink[`.tbl`]?
--
<br>.attn[A] Implicitly defined as `starwars` via the pipe (`starwars %>%`).

---
name: mutate-at-fun

For example, take the `log` of `height` and `mass`.

```{r, eval = F}
# Option 1a: Character vector of variable names
starwars %>% mutate_at(`.vars` = c("height", "mass"), `.funs` = log)
# Option 1b: Same, but don't name the arguments
starwars %>% mutate_at(c("height", "mass"), log)
```

.note[Above:] You don't have to name the arguments .orange[`.vars`] and .purple[`.funs`].

---

For example, take the `log` of `height` and `mass`.

```{r, eval = F}
# Option 2a: A "list" via the 'var' function
starwars %>% mutate_at(.vars = `vars(height, mass)`, .funs = log)
```

.note[Above:] You can "list" the target variables inside `vars()`.

---

For example, take the `log` of `height` and `mass`.

```{r, eval = F}
# Option 2b: Variable "span" inside 'var' 
starwars %>% mutate_at(.vars = `vars`(height`:`mass), .funs = log)
```

.note[Above:] .orange[.b[Inside] `vars()`], define a "span" of (sequential) variables with `:`.

---

For example, take the `log` of `height`-related variables.

```{r, eval = F}
# Option 2c: Matching names inside 'var'.
starwars %>% mutate_at(.vars = vars(`starts_with`("height")), .funs = log)
```

.note[Above:] .orange[.b[Inside] `vars()`], variables whose names start with `"height"`.

--

"Predicate" functions: `starts_with`, `ends_with`, `contains`, `matches`.

---
layout: true

# .mono[mutate] and family
## .mono[\_at]

---
name: mutate-at-ex

.attn[Q] How might we use `mutate_at` to help with missing values?

.note[Ex.] C-3PO and R2-D2 (among others) have `NA` for `hair_color`
--
,<br>but let's apply our technique for .b.orange[all `"_color"` variables].

--

We need two things:

.note[1\.] Which .b.orange[`.vars`] we will target.
--
```{r, eval = F}
.vars = vars(contains("_color"))
```

--

.note[2\.] The .b.purple[function(s)] (.purple[`.funs`]) for the mutation/transformation.
```{r}
# We'll write our own function but can rely upon 'replace_na'
na_to_other = function(d) replace_na(data = d, replace = "other")
```

---

The function `replace_na(data, replace)` comes from [`tidyr`](https://tidyr.tidyverse.org/reference/replace_na.html).

- .b.turquoise[`data`] can be a vector or data frame

- .b.slate[`replace`] is the replacement for those pesky `NA`s

.attn[Be careful:] Just because you .it[can] replace `NA`s doesn't mean you .it[should].

---
Let's put it all together now.

```{r, eval = F}
# Define the function to replace NAs with "other"
na_to_other = function(d) replace_na(data = d, replace = "other")
# Use mutate_at to transform "_color" variables
starwars %<>% mutate_at(
  .vars = vars(contains("_color")),
  .funs = na_to_other
)
# Check if it worked!
starwars %>% select(name, contains("_color"))
```

```{r, echo = F}
# Define the function to replace NAs with "other"
na_to_other = function(d) replace_na(data = d, replace = "other")
# Use mutate_at to transform "_color" variables
starwars %<>% mutate_at(
  .vars = vars(contains("_color")),
  .funs = na_to_other
)
# Check if it worked!
starwars %>% select(name, contains("_color")) %>% datatable(rownames = F, options = list(dom = 't', pageLength = 3))
```

---

.note[Note:] We don't actually have to define our own function.

`replace_na()` just wants to know what should `replace` the `NA`s.


```{r, eval = F}
# Use mutate_at to transform "_color" variables
starwars %<>% mutate_at(
  .vars = vars(contains("_color")),
  .funs = `replace_na`, 
  `replace = "other"`
)
```

```{r, echo = F}
# Use mutate_at to transform "_color" variables
tmp = dplyr::starwars %>% mutate_at(
  .vars = vars(contains("_color")),
  .funs = replace_na, 
  replace = "other"
)
# Check if it worked!
tmp %>% select(name, contains("_color")) %>% datatable(rownames = F, options = list(dom = 't', pageLength = 3))
```

---
layout: false
name: mutate-if-intro
# .mono[mutate] and family
## .mono[\_if]

`mutate`.neg-left.b.pink[`_if`] is very similar to `mutate`.neg-left.b.orange[`_at`]. 

- .b.pink[`_if`] uses .pink[logical statements] to select columns.

- .note[Recall:] .b.orange[`_at`] uses .orange[more direct statements] to select columns.

---
layout: true
# .mono[mutate] and family
## .mono[\_if]

`mutate_if()` wants three arguments:

1. .b.pink[`.tbl`] the dataset you're trying to mutate/transform (*e.g.*, `starwars`)
1. .b.orange[`.predicate`] logical statement to select variables (*e.g.*, `is_character`)
1. .b.purple[`.funs`] function(s) for the transformation (*e.g.*, `replace_na`)

---
name: mutate-if-fun

For example, let's .purple[change `NA`s to `"other"`] for .orange[all `character` variables].

```{r, eval = F}
# Replace NAs with "other" for character variables
starwars %>% mutate_if(
  .predicate = `is.character`,
  .funs = replace_na, replace = "other"
)
```

.note[Above:] .mono[R] finds all `character` variables and runs `replace_na` on them.

---
layout: false
name: mutate-if-ex
# .mono[mutate] and family
## .mono[\_if]

Let's .purple[change `NA`s to variables' means] for .orange[`numeric` variables].

.b.slate[Step 1] Create a function that replaces `NA`s with the mean:
```{r, eval = F}
# Function for mean imputation/replacement
mean_replace = function(d) replace_na(d, replace = mean(d, `na.rm = T`))
```
.note[Notice:] We need `na.rm = T` in `mean()` to calculate the mean when `NA`s exist.

--

.b.slate[Step 2] Use the new function `mean_replace()` with `mutate_if()`:
```{r, eval = F}
# Replace NAs with means for numeric variables
starwars %>% mutate_if(
  .predicate = is.numeric,
  .funs = mean_replace
)
```

---
layout: false
class: clear, middle 

What if we want to apply a function to .b.slate[all] of our columns?

---
layout: true
# .mono[mutate] and family
## .mono[\_all]

---
name: mutate-all-intro

`mutate_all(`.neg-left.b.pink[`.tbl`].neg-left[`,`].neg-left.b.purple[`.funs`].neg-left[`)`] does exactly what its name suggests. 

--

You supply the data (.b.pink[`.tbl`]). 

--

`mutate_all()` transforms .b[all] variables in .b.pink[`.tbl`] using .purple[.b[`.funs`] function(s)].

--

Now you just need a function that you want to apply to every variable...


---
layout: false
# The .mono[dplyr] family
## There's more!

Just as `mutate` has `mutate_at`, `mutate_if`, and `mutate_all`, 

- `transmute` ⮕ `transmute_at`, `transmute_if`, and `transmute_all`
- `select` ⮕ `select_at`, `select_if`, and `select_all`
- `filter` ⮕ `filter_at`, `filter_if`, and `filter_all`
- `summarize` ⮕ `summarize_at`, `summarize_if`, and `summarize_all`
- `group_by` ⮕ `group_by_at`, `group_by_if`, and `group_by_all`

--

.note[Note:] `dplyr` is working to replace the "scoped verbs" (`_if`, `_at`, `_all`) with the function [`across()`](https://dplyr.tidyverse.org/reference/across.html). But it's a bit less clean.

---
# Finishing up
## What's next?

Check out the .b.slate[wine reviews](https://www.kaggle.com/zynicide/wine-reviews) dataset. It has a lot of opportunities for data cleaning.

.note[Coming soon:]
- simulation, resampling, and cross validation
- [`tidymodels`](https://www.tidymodels.org), [`recipes`](https://recipes.tidymodels.org/index.html) (more [here](https://www.tidymodels.org/start/recipes/#recipe)), and [`parsnip`](https://parsnip.tidymodels.org/articles/parsnip_Intro.html)
- .it[Related:] [tidy modeling with .mono[R]](https://www.tmwr.org/models.html)

---

layout: false
# Table of contents

.col-left.small[

#### .mono[mutate]
- [Refresher](#mutate-refresh)
- [Pipes](#pipe-refresh)

#### .mono[mutate_at]
- [Intro](#mutate-at-intro)
- [Fundamentals](#mutate-at-fun)
- [Example](#mutate-at-ex)

]

.col-right.small[

#### .mono[mutate_if]
- [Intro](#mutate-if-intro)
- [Fundamentals](#mutate-if-fun)
- [Example](#mutate-if-ex)

#### .mono[mutate_all]
- [Intro](#mutate-all-intro)

]
